### 2010-12-04 paper.tex
Added latex paper

% !TEX TS-program = pdflatex
% !TEX encoding = UTF-8 Unicode

% This is a simple template for a LaTeX document using the "article" class.
% See "book", "report", "letter" for other types of document.

\documentclass[11pt]{article} % use larger type; default would be 10pt

\usepackage[utf8]{inputenc} % set input encoding (not needed with XeLaTeX)

%%% Examples of Article customizations
% These packages are optional, depending whether you want the features they provide.
% See the LaTeX Companion or other references for full information.

%%% PAGE DIMENSIONS
\usepackage{geometry} % to change the page dimensions
\geometry{a4paper} % or letterpaper (US) or a5paper or....
% \geometry{margins=2in} % for example, change the margins to 2 inches all round
% \geometry{landscape} % set up the page for landscape
%   read geometry.pdf for detailed page layout information

\usepackage{graphicx} % support the \includegraphics command and options

% \usepackage[parfill]{parskip} % Activate to begin paragraphs with an empty line rather than an indent

%%% PACKAGES
\usepackage{booktabs} % for much better looking tables
\usepackage{array} % for better arrays (eg matrices) in maths
\usepackage{paralist} % very flexible & customisable lists (eg. enumerate/itemize, etc.)
\usepackage{verbatim} % adds environment for commenting out blocks of text & for better verbatim
\usepackage{subfig} % make it possible to include more than one captioned figure/table in a single float
\usepackage{amstext}
% These packages are all incorporated in the memoir class to one degree or another...

%%% HEADERS & FOOTERS
\usepackage{fancyhdr} % This should be set AFTER setting up the page geometry
\pagestyle{fancy} % options: empty , plain , fancy
\renewcommand{\headrulewidth}{0pt} % customise the layout...
\lhead{}\chead{}\rhead{}
\lfoot{}\cfoot{\thepage}\rfoot{}

%%% SECTION TITLE APPEARANCE
\usepackage{sectsty}
\allsectionsfont{\sffamily\mdseries\upshape} % (See the fntguide.pdf for font help)
% (This matches ConTeXt defaults)

%%% ToC (table of contents) APPEARANCE
\usepackage[nottoc,notlof,notlot]{tocbibind} % Put the bibliography in the ToC
\usepackage[titles,subfigure]{tocloft} % Alter the style of the Table of Contents
\renewcommand{\cftsecfont}{\rmfamily\mdseries\upshape}
\renewcommand{\cftsecpagefont}{\rmfamily\mdseries\upshape} % No bold!

%%% END Article customizations

%%% The "real" document content comes below...

\title{Minimizing Time Required For Graduation}
\author{Ruyi Cai \and Matthew Fonda \and Michael Hughes}
%\date{} % Activate to display a given date or no date (if empty),
         % otherwise the current date is printed

\begin{document}
\maketitle

% Adds a table of contents. Comment out if we don't want this.
\tableofcontents
\pagebreak

\abstract{The Bachelor's Degree is designed to take four years to complete; however, many students fail to complete it this amount of time. One of the factors contributing to this is a lack of proper planning during the registration phase. In this paper we present an algorithm to assist students in constructing a schedule that works towards completing a degree in the shortest amount of time possible.}

\section{Introduction}
\subsection{Background}
Registration can be a stressful time for every student at the Univeristy of Washington. Without deliberate planning, students must decide which classes to take on a quarterly basis. The decision process involves considering many factors, such as quarterly class offerings, the amount of credits, the prerequisites of the classes, time conflicts, and how the choices in current quarter affect future quarters. The University of Washington registration system provides a schedule finder; however, it only takes into account time conflicts, and does not give advice on how to pick the most ideal schedule. The situation described above makes it difficult for students advising faculties. A student's graduation time may be extended because of his or her inability to take a class that has limited offerings during that particular year, for reasons such as not satisfying the prerequisites of the class. The advising office’s workload can also be increased due to the changing plans and majors of students and their need for repeated advising.

\subsection{Goal}
Our goal is to alleviate the problematic situation by creating a mathematical algorithm that accomplishes the following goal: given a student's current state and preferred credit hours per quarter, produce a class schedule that minimizes the number of quarters until graduation, eliminates time conflicts, and eliminates prerequiste conflicts.

\subsection{Importance}
The implementation of the algorithm as part of the school’s registration system will benefit the students by enabling them to map out their quarterly class schedules until graduation, and to preview their ideal schedule, the length of time until graduation, and the credit hours per quarter. This will lead to better decision making for what majors to pursue, and given the student's current state, which schedules are the most ideal in terms of saving time and resources. This will also help studentssave money, as they will be in school for fewer quarters. The university also benefits from the reduced labor cost and increased efficiency of the advising offices.

\section{Problem Description}
We focus on the classes offered within the Math department, although our algorithm and model can be easily extended to include over departments.

Every quarter the Math department offers a subset of classes from all of the available courses, each class is offered during a certain subset of available time periods. Since there are far more classes than time periods during the regular five-day week there are inevitably conflicts between courses that a student wants to take and courses that a student needs to take. Or worse, there are conflicts between two or more courses that a students needs to graduate. To graduate with a B.S. in Mathematics there are certain classes that are required and then there are elective requirements too. Additionally a student may wish to take some classes that are not required.
Adding to the problem is many classes cannot just be taken whenever they are available; instead, certain prerequisites must be satisfied first. In the case of math courses, there can be multiple different prerequisite sets that satisfy the requirements for a class.

Our goal is to determine a schedule of classes from some starting point to graduation that given the above properties minimizes the total number of quarters needed. Slightly more formally we would like to: \\
Minimize the total number of quarters needed to graduate subject to:
\begin{enumerate}
\item Only taking courses during quarters they are available
\item Only taking a course once
\item Never exceeding a preset number of credit hours per quarter
\item Never taking more than one class at a given time
\item Always satisfying a course's prerequisites prior to taking that course
\end{enumerate}

\subsection{Definitions}
We will now formalize some terms used in our paper.
\begin{itemize}
\item A {\it course} refers to any course offered by the University of Washington. A course has associated with it a name used in the University of Washington time schedule, for example {\it MATH309}. A course also has associated with it the quarters in which it is offered. Throughout this paper, we will refer to these using their time schedule abbreviates: {\it F} for Fall Quarter, {\it W} for Winter Quarter, {\it Sp} for Spring Quarter, and {\it S} for Summer Quarter. Each course is also worth a specific amount of credits.
\item We refer to a specific instance of a course as a {\it course offering}. Each course offering additionally has associated with it the time of the day at which the course meets, the quarter it is being offered, and the section number of the offering.
\end{itemize}

We must now also define some variables used throughout our paper.
\begin{itemize}
\item Let $S$ be the set of courses required to be taken in order to graduate. $S$ can be thought of as our goal state, and can also contain elective courses a student simply wishes to take.
\item Let $T$ be the set of courses a given student has already completed. If no courses have been taken, this set may be empty.
\item Let $U$ be the set of all courses a student could possibly ever take.
\item Let $P$ be a set of rules specifying course prerequistes.
\item Let $l$ be the maximum number of credits a student wishes to take during any given quarter.
\item Let $O$ be the set of all 50-minute time blocks between 8:30am and 7:30pm from Monday through Friday.
\end{itemize}

\subsection{Representing Prerequisites}
Initially, we hoped to represent course prerequistes using a directed acyclic graph, in which each node was a course, and edges between nodes $i$ and $j$ if and only if $i$ was a prerequisite for $j$.  This method was used by Pritsker et. al. and Xiangtong et. al. However, it does not work in our model, and courses can have a complicated set of multiple prerequisites. For example, {\it MATH327} requires both $MATH126$ and $MATH300$, or just $MATH136$. Let this graph be called $G_{prereq}$. Due to these complications, a $G_{prereq}$ alone will not suffice.

To solve this issues, we introduce a set of rules for each course. Each rule is a boolean expression represented in propositional logic in disjunctive normal form. For example, the rule assoicated with $MATH327$ is
$$
\text{{\it MATH327: (MATH126 AND MATH300) OR MATH136}}.
$$
As previously stated, we denote the set of all rules for all classes as $P$.

Furthermore, to determine if a given student is eligible to register for a course offering, we can evaluate the rule for the course, where each clause takes on the value true if a student has already taken the course, and false if he or she hasn't. If the rule evaluates to true, then the student is eligible to take the course; otherwise, the student is not.

\subsection{Initial Model and Tracability}
With a static schedule it would be possible to deterministically plan an entire schedule from start to finish, unfortunately we have neither the computational resources nor the luxury of a static schedule. From the student’s point of view, although there are some classes that are offered every quarter, senior electives are not, and sometimes it cannot be predicted more than one quarter in advance when a senior elective will be offered. Even given perfect information, where we know the possible class schedules for every quarter we face another problem; finding the best possible schedule reduces to enumerating possible class schedules for every quarter until the graduation requirements are satisfied. Given that we cannot predict in advance the conflicts between classes, we cannot even determine the branching factor of the enumeration tree nor its depth. Since we can only determine one quarter in advance, this suggests approaching the problem by using heuristics to pick the best possible classes each quarter.

Instead of trying to solve the complete problem we chose to use the rolling horizon approach used by Xiangtong et al \cite{xiangton:informs} when scheduling classes for airlines. Our problem already has a well defined horizon demarcation, the quarters. For the rest of the paper we focus on developing methods for attempting to optimally pick classes in each quarter separately.

\section{Rolling Horizons Approach}
We adapt a rolling horizons approach which uses a heuristic to pick the best possible schedule for any given quarter, given the classes offered during the quarter and the courses the student has previously taken. We use a greedy algorithm that functions on a per-quarter basis to accomplish this. The greedy algorithm greatly improves the tracability of course scheduling over an initially unknown number of quarters.

We break our algorithm down into two main phases: determining available course offering and selecting specific course offerings.

\subsection{Determining Available Course Offerings}
We need to select the classes we can currently take and then build conflict graphs for classes based on the times they are offered. In this section, we let $L$ be the initially empty set of course offerings a given student is eligible to take in the given quarter, $q$. \\

\subsubsection{Procedue}
We present the procedure for popluating the set $L$ as follows:
\begin{enumerate}
\item For each rule in $P$ we iterate over each proposition in the rule, $p$, and check if $p \in T$. If $p \in T$ then we say $p$ evaluates to true; otherwise, $p$ evaluates to false. We recall that each rule is associated with a given course. If the entire rule evaluates to true, then we add the course associated with it to $L$. Once this is complete, $L$ will consist of every course that could be taken based solely on prerequisites that we have fulfilled.
\item We now prune this set by removing all courses that are not offered in the current quarter. $L$ will necessarily now only contain classes that we can take given the current quarter and the classes we have already taken.
\item We now construct a confict graph, $G_{conflict}$, where every course offering in $L$ is a node, and there is an edge between two nodes $i$ and $j$ if and only if the time of course offering $i$ overlaps the time of course offering $j$.
\end{enumerate}

\subsubsection{Analysis}
The selection procedure involves a lot of set queries and iteration, but in asymptotic running time the algorithm is relatively quick. Let $n$ be the total number of classes ever available to take, or in other words, $n = \|U\|$. In the worst case, $P$ has a complex rule for all $n$ courses, and checking each rule in $P$ is linear in the number of propositions. However, we know {\it a priori} from analysis of our dataset that the largest rule has twelve propositions, so we can bound the rule check time by a constant, $C$, and therefore step one can be done in $O(Cn)$ time.

Continuing to step two, the worst case is when $L$ = $Q$. Checking whether a course is offered in a given quarter can be done in constant time, and we have $n$ courses to check, therefore step two can also be done in $O(Cn)$ time.

Finally, the last step is once again bounded by a constant and $n$. Considering courses between 8:30am and 7:30pm, we have nine fifty-minute time periods, and five days of the week, and therefore we have 45 time slots to check in our conflict graph, again yielding $O(Cn)$ runtime.

Since we are dealing with asymptotic running time we will drop the constant terms and combining all resulsts yields an overall runtime of $O(n)$.

\subsection{Selection of Quarterly Classes}
\subsubsection{Selection Prodcedure}
At this point in time, we know which course offerings are available for the given student to register for, and we must now pick the most optimal ones. There a several ways we could view the problem of determining which classes to take each quarter. We will initially encode the class selection as a binary integer programming problem in spirit of the methods presented by Dinkel et al \cite{dinel:scheduling}. and Pritsker et al \cite{prisker:informs}. At this point we have the set $L$ containing possible classes we that we need to take.

\subsubsection{Heuristic Value Function}
In order to determine which classes should be picked in a given schedule, we use a heuristic function which assigns each course 
offering a value. The higher the value of this function for a given course offering is, the more likely the given course offering 
will be placed in the schedule. Recalling that our goal is to minimize the total number of quarters taken (and implicitly, 
maximize the total number of courses taken per quarter), our heuristic function should clearly give higher value to course 
offerings that will help us achieve this goal. One component to the heuristic function then must be the number of dependent 
courses a given course offering has. If a course satisfies many prerequisties, then we know that there are many courses that 
we cannot complete without it, so we want our heuristic function to favor it.

Additionally, we know that not every course is offered every quarter. If a course is not offered until the following year, we 
could potentially add an entire year to the schedule. This is a situation we hope to avoid. Our heuristic function must 
therefore consider how often and how far in the future a given course is offered. For example, if we are deciding between 
two course offering that are identical except one is ofference once a year, and the other is offered three times a year, we 
want to take the course only offered once a year, as we won't have the oppurtunity to take it for another entire year.

We these considerations in mind, we begin to develop our heuristic function. In order to define our heuristic function, we 
must first define some auxilary functions on course offering.
\begin{itemize}
\item Let $TimeFromNow(c, q)$ be difference in quarters between the current quarter, $q$, and the next quarter a given course offering, $c$, is offered. As there are four quarters, the range of this function is $\{0, 1, 2, 3\}$.
\item Let $Quarters(c)$ be the number of quarters per year a given course, $c$, is offered. As there are four quarters, the range of this function is $\{0, 1, 2, 3, 4\}$.
\item Let $Dependents(c)$ be the number of course that either directly or indirectly require a given course, $c$. In our final 
draft, we will present a diagram illustrating this value more clearly.
\end{itemize}

\includegraphics{more_prereq_tree}

With these functions defined, we can now define our heuristic function:
\begin{equation}
Value(c) = 1 + \alpha * Dependents(c) + \beta * TimeFromNow(c, q) + \gamma * Quarters(c)
\end{equation}
Where $c$ is a given course offering, $q$ is the current quarter, and $\alpha$, $\beta$, and $\gamma$ are real numbers, giving a 
weight to piece of the heuristic. The scalars are determined experimentally.

\subsubsection{Formulation as Binary Integer Program}
We can now formulate our problem as a binary integer linear programming probelm. Before stating the linear program, we must first define some variables used in it: \\
\begin{itemize}
\item $
  X_i = \left\{
     \begin{array}{lr}
       1 \text{ if we take } X_i\\
       0 \text{ if we don't take } X_i
     \end{array}
   \right. $
\item For all $t \in O$, we let $t = 1$.
\item Let $m = \|L\|$.
\end{itemize}
We now state our problem as follows:\\
Maximize:
$$
\text{Value of Quarter} = \sum_{i=1}^m Value(c_i) * X_i
$$
Subject To:
$$
\sum_{i=1}^m Weight(c_i) * X_i \leq l,
$$
$$
\forall t_j \in O, \sum_{i=1}^m t_j * X_i \leq 1,
$$
$$
X_i \in \{0, 1\}
$$
The linear program can be interpreted as follows:
\begin{itemize}
\item The value of a quarter is the total number of courses taken that quarter.
\item In each quarter, the total number of credits cannot exceed the specified limit.
\item Only one course can be taken in one particular time slot
\item A course can only be taken or not taken. In other words, decision variables are binary.
\item Each non-zero descision variable in the optimal solution represents a class we will take.
\end{itemize}

It is worth nothing that all of the t variables are essentially dummy variables that ensure the correct number of equations of type 2. Each of the equations based on eqn. 2 ensure that we only take one class per time period per day. Also this linear program is ‘nice’ in that it always has feasible origin (don’t take any classes), and that the simplest formulation is already in standard form.

As noted in the reference literature this is actually a (strongly) NP-complete problem as described \cite{pferschy:kcg}. For the rest of this section we consider an alternative formulation of class selection as a knapsack problem with conflict graphs and suggest the use a technique developed by Yamada et. al. \cite{yamada:heuristic}.

We begin by simplifying $G_{conflict}$ to be a fully connected graph by adding dummy nodes $v'$ and insert edges that connect $v'$ to every other node in the graph. To ensure that $v'$ does not alter the final optimal solution, we assign it the following properties:
\begin{itemize}
\item $Weight(v') = l+1$
\item $Value(v') = 0$
\end{itemize}
As the weight of our dummy node is greater than our credit limit, and it's value is zero, we know that it will never appear in the optimal solution. With the above reasoning we show for our case of class schedule that Pferschy and Schauer's \cite{pferschy:kcg} claim that adding a dummy node without loss of generality holds. We can now conveniently write the class picking problem as the knapsack with conflict graphs problem (KCG): \\
Maximize:
$$
\text{Value of Quarter} = \sum_{i=1}^m Value(c_i) * X_i
$$
Subject To:
$$
\sum_{i=1}^m Weight(c_i) * X_i \leq l,
$$
$$
X_i + X_j \leq 1 \text{ for all connected vertices } i, j \in G_{conflict},
$$
$$
X_i \in \{0, 1\}
$$
It is worth noting that for relatively small versions of this problem, in the low hundreds of variables it is possible to solve this problem exactly using modern LP solving software \cite{yamada:heuristic}. In our case, just solving for the optimal classes for a single major fits well into the scope of a few hundred (or even just tens of) variables. For multiple majors or to consider class selection for distribution and elective requirements, however the number of classes balloons significantly and the KCG problem becomes much harder. For this version of the problem we suggest using the techniques used by Yamada et al \cite{yamada:heuristic}, which we briefly outline below.

\subsubsection{Relaxation and Implicit Enumeration}
\begin{enumerate}
\item Develop upper bound on value by solving a continuous version knapsack problem generated by a Lagrangean relaxation of our problem. Let $X_i$ be a continuous variable with $0 \leq X_i \leq 1$, and shift the time constraints into the objective function by using a penalty of $\lambda$ when the constraint is violated.
\item Use the upper bound developed above with the pruning rules suggested by Yamada et al \cite{yamada:heuristic} to do implicit enumeration of the possible choices.
\end{enumerate}

In Yamada et al’s \cite{yamada:heuristic} numerical runtime analysis they found this algorithm ran very favorably against solving the same problem using popular linear programming solvers. There is one problem that we should briefly address, because the number of periods of available classes per week is relatively small the conflict rate for a large number of class is relatively higher than Yamada et al’s 2\% maximal conflict rate for their experiments. Our data has a significant mitigating factor though, knapsack problems in general are exponential in terms of the number of choices and the total weight capacity. Realistically the total weight capacity is going to be 20 credits or less which is between 1\% and 2\% of the figures used by Yamada \cite{yamada:heuristic} and also our total number of choices is in many cases (particular just the classes in a particular major) less than 200. Giving the size and type of data in our set the higher conflict rate is less of a concern and Yamada et al’s algorithm should still be efficient.

\section{Computational Results}
In this section we will outline results obtained by running computer simulations of schedule picking. We will computationally examine what parameters in our heuristic function produce the most desirable schedules.

\bibliography{references}{}
\bibliographystyle{plain}
\end{document}